Presolving for\\ Mixed-Integer Semidefinite Optimization

### 2020-12-19 tex/MISDP-Presolving/MISDPpresolving.tex
handle linear case of feasibility set

\subsubsection{Complete Description of Feasible Set}

@@

\paragraph{Linear Case}

It might be instrumential to look at the linear analogue:
\[
  S \define \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i =
  1,\; \sum_{i=1}^n z_i \leq k,\; 0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
\]
Both the polytopes
$\{x \in \R^n \suchthat \ones\T x = 1,\; 0 \leq x \leq z\}$ for fixed
$z \in \{0,1\}^n$ and
$\{z \in \R^n \suchthat \ones\T z \leq k,\; x \leq z\}$ for fixed
$x \in \{0,1\}^n$ are integral.

\begin{question}
  Is the above description integral?
\end{question}

>>>>

% -------------------------------------------------------------------------
\subsubsection{Complete Description of Feasible Set}

@@

% -------------------------------------------------------------------------
\paragraph{Linear Case}

It might be instrumential to look at the following linear analogue:
\[
S^k \define \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i =
1,\; \sum_{i=1}^n z_i = k,\; 0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
\]
As before, $S^k_I$ denotes the integer hull
$S^k_I \define \conv\big(S^k \cap (\R^n \times \{0,1\}^n)\big)$.

\begin{theorem}
  $S^k = S^k_I$, i.e., $S^k$ is integral.
\end{theorem}

\begin{proof}
  The constraint matrix of $S^k$ with respect to the equations and the
  inequalities $x - z \leq 0$ can be written as follows:
  \[
    M \define
    \begin{pmatrix}
      1 & \dots & 1 & 0 & \dots & 0 \\
      0 & \dots & 0 & 1 & \dots & 1 \\
      1 &       &   & -1 &      &   \\
        & \ddots&   &    & \ddots & \\
      0 &       & 1 & 0 &      & -1
    \end{pmatrix}.
  \]
  This matrix is totally unimodular. This can be seen by multiplying the
  first row with $-1$. Then we have exactly one $+1$ and $-1$ in each
  column. Thus, $M$ is the node-arc indicence matrix of a directed graph
  and hence TU. To take care of the variables bounds, one would add
  identity matrices to $M$, which do not change the property of being TU.
\end{proof}

As a consequence, when solving
\[
  \min\, \{ c\T x + d\T z \suchthat (x,z) \in S^k \}
\]
one can even chose the $x$-variables to be integral, i.e., in 0 or 1, in an
optimal solution.

\begin{remark}
  Aghezzaf and Wolsey~\cite{AghW92} show the following: Let
  $T \subseteq \R^p_+ \times \{0,1\}^n$, $P = \conv(T)$, and
  \[
  P_k \define \{ (x,z) \in P \suchthat \sum_{i=1}^n z_i = k\}.
  \]
  For $k = 0, \dots, n$ define
  \[
  g_k \define \min\,\{c\T x + d\T z \suchthat (x,z) \in P_k,\; z \in
  \{0,1\}^n\}.
  \]
  Then they show that
  \[
  \min\,\{c\T x + d\T z \suchthat (x,z) \in P_k\}
  \]
  has an optimal solution in $T$ if and only if $g_{k+1} - g_k$ is
  non-decreasing for $k = 1, \dots, n - 1$.
  % Alternatively, if for some $\mu$
  % \[
  % \min\, \{c\T x + d\T z - \mu \sum_{i=1}^n z_i \suchthat (x,z) \in P\}
  % \]
  % has integral optimal solutions $(x^1,z^1)$ and $(x^2,z^2)$ with
  % $\ones\T z^1 = k_1$ and $\ones\T z^2 = k_2$ with $k_1 \leq k_2 - 2$, then
  % there exists an alternative integral optimal solution with $\ones\T z = k$ and
  % $k_1 < k < k_2$. Note that one can integrate the $\mu$-term into $d$.

  In our situation we have
  \[
  T = \{(x,z) \in \R_+^n \times \{0,1\}^n \suchthat \sum_{i=1}^n x_i = 1,\;
  0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
  \]
  The convex hull of $T$ is simply obtained by relaxing $z$ to be in $[0,1]^n$:
  \[
  P = \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i = 1,\;
  0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
  \]
  Thus, $P_k = S^k$.

  Then Aghezzaf and Wolsey consider lot sizing problems for given demands $D_1,
  \dots, D_n$:
  \begin{align*}
    \min\quad & c\T x + d\T z\\
    & \sum_{i=1}^t x_i \geq \sum_{i=1}^t D_i\quad \forall i = 1, \dots, n - 1,\\
    & \sum_{i=1}^n x_i = \sum_{i=1}^n D_i,\\
    & 0 \leq x \leq M z,\\
    & z \in \{0,1\}^n.
  \end{align*}
  If $D_i = 0$, $i = 1, \dots, n-1$, $D_i = 1$, and $M = 1$, we obtain our problem.

  Aghezzaf and Wolsey show that the optimum is attained in an integral
  point for each $k$ if $c_1 \geq c_2 \geq \dots \geq c_n$. But the proof
  is not completely clear.
\end{remark}
